---
title: "Basic Viz"
author: "Selena Buttery"
date: "2025-04-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)


library(vembedr)
```

```{r}
sdr_data <- read_csv(here("data/SDR-2023-Data.csv"))

sdr_data <- sdr_data %>% 
  clean_names()
```

```{r}
ggplot(data = sdr_data, aes(x= goal_3_score, y = goal_4_score)) +
  geom_point()
```

```{r}
ggplot(data = sdr_data, aes(x= goal_3_score, y = goal_4_score, color = regions_used_for_the_sdr)) +
  geom_point()
```

```{r}
ggplot(data = sdr_data, aes(x= goal_3_score, y = goal_4_score, color = goal_5_score)) +
  geom_point()
```

Challenge 1

```{r}
ggplot(data = sdr_data, aes(x= goal_4_score, y = goal_7_score, color = goal_12_score)) +
  geom_point()
```

Explanation:

-   We do see in this graph that as all the goals increase in score, they all do it together except there is a little bit of an outlier when we see the darker dot, it is high for goal 4 but not goal 7 or goal 12 which is interesting.

```{r}
ggplot(sdr_data, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity")
```

```{r}
lac_sdr_data <- sdr_data %>% 
  filter(regions_used_for_the_sdr == "LAC")

ggplot(lac_sdr_data, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity") 
```

```{r}
ggplot(sdr_data, aes(x= goal_3_score, y = country)) +
  geom_bar(stat = "identity") +
  facet_wrap(~regions_used_for_the_sdr, scales = "free_y") + # faceting for regions
  theme(axis.text.y = element_text(size = 4)) 
```

Challenge 2:

```{r}
MENA_sdr_data <- sdr_data %>% 
  filter(regions_used_for_the_sdr == "MENA")


ggplot(MENA_sdr_data, aes(x= goal_12_score, y = country)) +
  geom_bar(stat = "identity") 
```

```{r}
ggplot(sdr_data, aes(x = goal_4_score)) +
  geom_histogram()
```

```{r}
ggplot(sdr_data, aes(x = goal_4_score, fill=regions_used_for_the_sdr)) + geom_histogram(bins=33, color="black")

```

Challenge 3:

```{r}
ggplot(sdr_data, aes(x = goal_12_score, fill=regions_used_for_the_sdr)) + geom_histogram(bins=33, color="black")
```

 Countries that fall in the higher end of the Goal 12 spectrum belong to the Sub-Saharan Africa region, while countries in the OECD lie towards the lower end. This is opposite of SDG goal number 4 shown above for Quality Education.
